### 2021-04-19 docs/pap/compromise-punishment-pap.Rmd
add changes to pap

---
title: 'Pre-Analysis Plan'
author: '16 April 2021'
output: 
  github_document:
    toc: yes
    number_sections: true
  bookdown::pdf_book:
    base_format: rticles::jss_article
editor_options: 
  chunk_output_type: console
bibliography: references.bib
---

# Hypothesis

# Research Desing and Protocol

## Sample
I will conduct this survey experiment in the Netherlands in April 2021.
Our sample, recruited through [KiesKompas](https://www.kieskompas.nl/en/), will consist of 2,000 participants (based on the power analysis presented below) of 18 years and older.
The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the _Vrije Universiteit Amsterdam_ (see the approval [here](../EthicalApproval.pdf).
To ensure good quality of our data, two attention checks (discussed in more detail below) are included.
Each respondent failing the attention check will be excluded and replaced with another 'good' response.

## Experimental Protocol
The study is conducted online and in Dutch.
Participants are told that they are taking part in a survey to get an overview of how Dutch people form their views on politics.
After reading an informed consent message participants are forwarded to the main questionnaire (or the survey will be terminated if they do not agree to the consent form).

First, participants complete a set of background variables on their stances on the political issues used in the experiment (i.e. corona-measures, defense, education, immigration) and on their attitude towards women in politics -- the full codeboek can be viewed [here](#).
The pre-treatment block ends with one of the two attention checks included in this survey.
When participants fail this attention check, a warning appears asking them to read the question again
carefully  and  to  answer  again.
Only when they have answered it correctly, they enter the first round of the experiment.
After each round of the experiment, some filler questions are asked about the respondents demographics.
The stimuli in the experiment are news messages in the same style as the Dutch news website [nu.nl](https://www.nu.nl/).
In these news messages, we manipulate: a) the _gender_ of the politician (male vs. female); b) the _migration background_ of the politician (based on a Arabic sounding name (Rachid(a) Amezian) or a native Dutch sounding name (Karel/Karin van der Kleijn)); and c) whether the politician _striked a compromise or not_.
This creates a full `2*2*2` factorial experiment with four rounds.
Every round, the news message covers a new issue.
The issues are: "defense" (increase cooperation of the defense units), "immigration" (immigration stop in western countries), "education" (expansion creative elementary school activities), "COVID-19" (emergency legislation to combat COVID-19).
For every round, the politicians name, and thereby gender and immigration background, are manipulated in the text and in the picture as well as the decision to strike a compromise.
For an illustration of the stimulus material, see the Figure below.
All manipulated elements are within the blue boxes.

```{r ssa-logo, fig.align = 'center', out.width = "75%", fig.cap = "The new SSA logo, which is actually a scatterplot, which is super neat!", echo = FALSE}
knitr::include_graphics(here::here("docs/pap/FMPdefensie_annotated.png"))
```

The text of the treatment is a Dutch translation of the stimulus material developed by @bauer2017punished. 
The authors have pre-tested the text of the stimuli material several times [see the Web Appendix of @bauer2017punished].
The photo's of their stimuli material can unfortunately not be used, because a) @bauer2017punished only investigate the difference between men and women politicians, using solely photo''s of caucasians; and b) the American flag was present in their material.
Instead, I have used photo's from local politicians of the city of The Hague that are not know to the general public. 
The men politicians are in real life representatives of the [liberal-progressive party D66](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/boutkabout-a.-abid-fractievertegenwoordiger-d66.htm) (man with immigrant background) and of the [christian democratic party CDA](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/groot-de-d.p.-daniel-fractievertegenwoordiger-cda.htm) (man without immigrant background).
The women politicians are in real life representatives of a [right-wing populist party _Hart voor Deen Haag/ Groep de Mos_](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/davituliani-n.-nino-raadslid-hart-voor-den-haaggroep-de-mos.htm) (woman with migration background) and of the [social democratic party PvdA](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/holman-j.-janneke-raadslid-pvda.htm) (woman without immigrant background).
Thereby we have a spread over the ideological different parties present in Dutch politics.
The names for the politicians without immigration background are _Karel_ (man) and _Karin_ (woman) _van der Kleijn_, which are typical native Dutch names.
For the politicians with an immigration background, we picked _Rachid_ (man) and _Rachida_ (woman) _Amezian_.

## Power Analysis
As detailed in [Section 4](#analysis), I conduct an OLS regression within a multiverse approach [@simonsohn2019specification,@steegen2016increasing] using the four post-treatment questions on the politician's a) leadership evaluaton (scale); b) general favorability; c) representative quality; and d) career perspective as dependent variable and the three manipulations (gender, migration background and whether or not the politician is willing to strike a compromise ) as independent variables.
Each hypothesis is tested separately for all four issues.

We will test each hypothesis separately for all
four issues.  We used the R package "DeclareDesign" (Blair et al., 2019) for our power
analysis 1, assuming a small effect of b= 0.2 (Figure 2, yellow line).  As our hypotheses are  directional,  we  will run  one-tailed  tests  with  an  alpha  of  .05.   Our  power  analysis shows that testing hypotheses 1 (a, b and c) and 3 (a and b), requires a sample size of 1,200 participants (x-axis) to reach power of 0.95 (black dashed line, Figure 2, panel A). Note that if the effect size b is bigger then .2 { i.e., .3, .4, .5 or .6 { smaller sample sizes
are sufficient to reach power of 0.95. To test Hypotheses 2 (Anger frame) and 4 (Anxiety Frame), we can detect a small effect b = 0.2 (yellow line) with a power of .8 (gray dashed line) and a one-tailed test of signficance at alpha .05 with a sample of 1,500 participants (Figure 2, Panel B). At 1,800 participants, with a small effect and one-tailed test of significance at .05, we reach power of .88.  A sample of 1.800 participants therefore gives us suficient power to test Hypotheses 1 and 3 but a probability of 12% for a Type II error remains when testing
Hypotheses 2 and 4.  We will therefore test Hypotheses 2 and 4 first by issue and second,
by pooling our data across issues.  This will give us sufficient power to test H2 and 4.

# Measures

## Dependent Variables

## Control Variables

## Attention Checks

## Exclusion Criteria

# Analysis

## Hypothesis 1

## Hypothesis 2

## Hypothesis 3

# Stimulus Material