### 2021-05-03 src/data-processing/prep_data.Rmd

---
title: 'Prepare Data'
author: 
output: 
  github_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center",
                      fig.path=('../../report/figures/'))
library(printr)
options(digits = 2)
library(rmarkdown)
```

# Scripts
- [Required Packages &amp; Reproducibility](#required-packages-&amp;-reproducibility)
- [Tidy Data](#tidy-data)
- [Save Data for Analysis](#save-data-for-analysis)
- [Check Missing Values](#check-missing-values)
- [Visualization of Data](#visualization-of-data)
  - [Dependent Variable](#dependent-variable)
  - [Treatment Conditions](#treatment-conditions)
  - [Correlations Matrix](#correlations-matrix)

## Required Packages &amp; Reproducibility
```{r "env", message=FALSE, warning=F}
rm(list=ls())
source(here::here("src/lib/functions.R"))
renv::snapshot()
```

## Tidy Data
This code chuck downloads the data from Qualtrics using the API and cleans the raw data.

```{r "get data from Qualtrics", results="hide"}
source(here("src/data-processing/tidy_data.R"))
```

## Check Missing Values

We employ the following criteria:

- If 10% or less of the values on the dimension are missing, then we re-code the missing values to the overall mean.
- If 11% or more of the values on the dimension are missing, then we re-code the missing values to a constant (for instance 0) and include a dummy variable indicating whether the response on the covariate was missing or not.

```{r "check missings"}

```

## Save Data for Analysis
```{r "save data",eval=FALSE}
save(d, file = here("data/intermediate/cleaned_data.RData"))
```

## Visualization of Data

### Dependent Variable
```{r "Dependent Variable", echo=F}
load(here("data/intermediate/cleaned_data.RData"))
d %>%
  mutate(`PT-D1` = round((`PT-D1_1` + `PT-D1_2` + `PT-D1_3` + `PT-D1_4` + `PT-D1_5` + `PT-D1_6`)/6, 0),
         `PT-E1` = round((`PT-E1_1` + `PT-E1_2` + `PT-E1_3` + `PT-E1_4` + `PT-E1_5` + `PT-E1_6`)/6, 0),
         `PT-C1` = round((`PT-C1_1` + `PT-C1_2` + `PT-C1_3` + `PT-C1_4` + `PT-C1_5` + `PT-C1_6`)/6, 0),
         `PT-I1` = round((`PT-I1_1` + `PT-I1_2` + `PT-I1_3` + `PT-I1_4` + `PT-I1_5` + `PT-I1_6`)/6, 0),
         PreT5 = round((PreT5_1 + PreT5_2 + PreT5_3 + PreT5_4)/4,0),
         PreT7 = round((PreT7_1 + PreT7_2 + PreT7_3 + PreT7_4 + PreT7_5 + PreT7_6 + PreT7_7)/7,0)) %>%
  pivot_longer(cols = c(`name-D`, `name-I`, `name-E`, `name-C`),
               values_to = "name",
               names_to = "issue") %>%
  drop_na(name) %>%
  mutate(issue = recode(issue, 
                        `name-C` = "COVID-19",
                        `name-D` = "Defense",
                        `name-E` = "Education",
                        `name-I` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D1`, `PT-E1`, `PT-C1`, `PT-I1`),
               values_to = "PT1",
               names_to = "issue1") %>%
  drop_na(PT1) %>%
  mutate(issue1 = recode(issue1, 
                        `PT-C1` = "COVID-19",
                        `PT-D1` = "Defense",
                        `PT-E1` = "Education",
                        `PT-I1` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D2`, `PT-E2`, `PT-C2`, `PT-I2`),
               values_to = "PT2",
               names_to = "issue2") %>%
  mutate(issue2 = recode(issue2, 
                        `PT-C2` = "COVID-19",
                        `PT-D2` = "Defense",
                        `PT-E2` = "Education",
                        `PT-I2` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D3`, `PT-E3`, `PT-C3`, `PT-I3`),
               values_to = "PT3",
               names_to = "issue3") %>%
  mutate(issue3 = recode(issue3, 
                         `PT-C3` = "COVID-19",
                         `PT-D3` = "Defense",
                         `PT-E3` = "Education",
                         `PT-I3` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D4`, `PT-E4`, `PT-C4`, `PT-I4`),
               values_to = "PT4",
               names_to = "issue4") %>%
  mutate(issue4 = recode(issue4, 
                         `PT-C4` = "COVID-19",
                         `PT-D4` = "Defense",
                         `PT-E4` = "Education",
                         `PT-I4` = "Immigration"),
        check = if_else(issue == issue1 & issue == issue2 & issue == issue3 & issue == issue4, 1,0)) %>%
  filter(check == 1) %>%
  select(PT1, PT2, PT3, PT4, issue = issue1) %>%
  pivot_longer(cols = PT1:PT4,
               names_to = "dvs") %>%
  mutate(dvs = recode(dvs,
                       `PT1` = "DV: Trait Evaluation",
                       `PT2` = "DV: Favorability",
                       `PT3` = "DV: Representation",
                       `PT4` = "DV: Career Prospects")) %>%
  group_by(issue, dvs) %>%
  summarise(means = mean(value, na.rm=T),
            stdev = sd(value, na.rm=T)) %>%
  mutate(lower = means - (1.56 * stdev),
         upper = means + (1.56 * stdev)) %>%
  ggplot(aes(x = dvs, y = means, colour = issue,
             ymin = lower, ymax = upper, label = issue)) +
  geom_point(position = position_dodge(.2)) + 
  geom_errorbar(position = position_dodge(.2), width = 0) +
  geom_label_repel(nudge_x = 0.35,nudge_y = 0.5, size = 4) + 
  labs(y = "0 (negative) - 10 (positive)", 
       x = "", title = "Evaluation of Politicians") +
  geom_hline(yintercept = 5, linetype = "dotted", color = "darkgrey") +
  theme_minimal() +
  scale_color_manual(values = fig_cols) +
  coord_flip() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="none",
        legend.title = element_blank()) 
```

### Treatment Conditions
```{r "Independent Variables", echo=F, message=FALSE, warning=FALSE,fig.width=10, fig.height=10}

d %>%
  mutate(`PT-D1` = round((`PT-D1_1` + `PT-D1_2` + `PT-D1_3` + `PT-D1_4` + `PT-D1_5` + `PT-D1_6`)/6, 0),
         `PT-E1` = round((`PT-E1_1` + `PT-E1_2` + `PT-E1_3` + `PT-E1_4` + `PT-E1_5` + `PT-E1_6`)/6, 0),
         `PT-C1` = round((`PT-C1_1` + `PT-C1_2` + `PT-C1_3` + `PT-C1_4` + `PT-C1_5` + `PT-C1_6`)/6, 0),
         `PT-I1` = round((`PT-I1_1` + `PT-I1_2` + `PT-I1_3` + `PT-I1_4` + `PT-I1_5` + `PT-I1_6`)/6, 0),
         PreT5 = round((PreT5_1 + PreT5_2 + PreT5_3 + PreT5_4)/4,0),
         PreT7 = round((PreT7_1 + PreT7_2 + PreT7_3 + PreT7_4 + PreT7_5 + PreT7_6 + PreT7_7)/7,0)) %>%
  pivot_longer(cols = c(`name-D`, `name-I`, `name-E`, `name-C`),
               values_to = "name",
               names_to = "issue") %>%
  drop_na(name) %>%
  mutate(issue = recode(issue, 
                        `name-C` = "COVID-19",
                        `name-D` = "Defense",
                        `name-E` = "Education",
                        `name-I` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D1`, `PT-E1`, `PT-C1`, `PT-I1`),
               values_to = "PT1",
               names_to = "issue1") %>%
  drop_na(PT1) %>%
  mutate(issue1 = recode(issue1, 
                        `PT-C1` = "COVID-19",
                        `PT-D1` = "Defense",
                        `PT-E1` = "Education",
                        `PT-I1` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D2`, `PT-E2`, `PT-C2`, `PT-I2`),
               values_to = "PT2",
               names_to = "issue2") %>%
  mutate(issue2 = recode(issue2, 
                        `PT-C2` = "COVID-19",
                        `PT-D2` = "Defense",
                        `PT-E2` = "Education",
                        `PT-I2` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D3`, `PT-E3`, `PT-C3`, `PT-I3`),
               values_to = "PT3",
               names_to = "issue3") %>%
  mutate(issue3 = recode(issue3, 
                         `PT-C3` = "COVID-19",
                         `PT-D3` = "Defense",
                         `PT-E3` = "Education",
                         `PT-I3` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D4`, `PT-E4`, `PT-C4`, `PT-I4`),
               values_to = "PT4",
               names_to = "issue4") %>%
  mutate(issue4 = recode(issue4, 
                         `PT-C4` = "COVID-19",
                         `PT-D4` = "Defense",
                         `PT-E4` = "Education",
                         `PT-I4` = "Immigration"),
        check = if_else(issue == issue1 & issue == issue2 & issue == issue3 & issue == issue4, 1,0)) %>%
  filter(check == 1) %>%
  mutate(gender = recode(name,
                         `Karel van der Kleijn` = 0,
                         `Karin van der Kleijn` = 1,
                         `Rachid Amezian` = 0,
                         `Rachida Amezian` = 1),
         immigration = recode(name,
                              `Karel van der Kleijn` = 0,
                              `Karin van der Kleijn` = 0,
                              `Rachid Amezian` = 1,
                              `Rachida Amezian` = 1),
         intersection = recode(name,
                               `Karel van der Kleijn` = 0,
                               `Karin van der Kleijn` = 0,
                               `Rachid Amezian` = 0,
                               `Rachida Amezian` = 1),
         compromise = if_else(compromis == "geen compromis", "Compromise", "No Compromise"))%>%
  unite("treatment", c("name", "compromise"), sep = "\n ") %>%
  select(treatment, issue) %>%
  group_by(issue, treatment) %>%
  summarise(n = n()) %>%
  mutate(perc = round(n/sum(n),2),
         perc2 = paste0(perc*100, "%")) %>%
  ggplot(aes(x = perc, y = issue, fill = treatment, label = perc2)) +
  geom_col(position = 'fill', width=0.75) + 
  geom_text(position = position_fill(vjust=0.5), color = "white", fontface = "bold") +
  labs(y = "", x = "", title = "Distribution of Treatments") +
  scale_x_continuous(labels=scales::percent) +
  theme_minimal() +
  scale_fill_manual(values = fig_cols) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom",
        legend.title = element_blank()) +
  guides(fill=guide_legend(ncol=4,byrow=TRUE))
```

### Correlations Matrix
```{r "Correlations Matrix", echo=F, message=T, warning=T, fig.width=15, fig.height=15}
d %>%
  mutate(`PT-D1` = round((`PT-D1_1` + `PT-D1_2` + `PT-D1_3` + `PT-D1_4` + `PT-D1_5` + `PT-D1_6`)/6, 0),
         `PT-E1` = round((`PT-E1_1` + `PT-E1_2` + `PT-E1_3` + `PT-E1_4` + `PT-E1_5` + `PT-E1_6`)/6, 0),
         `PT-C1` = round((`PT-C1_1` + `PT-C1_2` + `PT-C1_3` + `PT-C1_4` + `PT-C1_5` + `PT-C1_6`)/6, 0),
         `PT-I1` = round((`PT-I1_1` + `PT-I1_2` + `PT-I1_3` + `PT-I1_4` + `PT-I1_5` + `PT-I1_6`)/6, 0),
         PreT5 = round((PreT5_1 + PreT5_2 + PreT5_3 + PreT5_4)/4,0),
         PreT7 = round((PreT7_1 + PreT7_2 + PreT7_3 + PreT7_4 + PreT7_5 + PreT7_6 + PreT7_7)/7,0)) %>%
  pivot_longer(cols = c(`name-D`, `name-I`, `name-E`, `name-C`),
               values_to = "name",
               names_to = "issue") %>%
  drop_na(name) %>%
  mutate(issue = recode(issue, 
                        `name-C` = "COVID-19",
                        `name-D` = "Defense",
                        `name-E` = "Education",
                        `name-I` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D1`, `PT-E1`, `PT-C1`, `PT-I1`),
               values_to = "PT1",
               names_to = "issue1") %>%
  drop_na(PT1) %>%
  mutate(issue1 = recode(issue1, 
                        `PT-C1` = "COVID-19",
                        `PT-D1` = "Defense",
                        `PT-E1` = "Education",
                        `PT-I1` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D2`, `PT-E2`, `PT-C2`, `PT-I2`),
               values_to = "PT2",
               names_to = "issue2") %>%
  mutate(issue2 = recode(issue2, 
                        `PT-C2` = "COVID-19",
                        `PT-D2` = "Defense",
                        `PT-E2` = "Education",
                        `PT-I2` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D3`, `PT-E3`, `PT-C3`, `PT-I3`),
               values_to = "PT3",
               names_to = "issue3") %>%
  mutate(issue3 = recode(issue3, 
                         `PT-C3` = "COVID-19",
                         `PT-D3` = "Defense",
                         `PT-E3` = "Education",
                         `PT-I3` = "Immigration")) %>%
  pivot_longer(cols = c(`PT-D4`, `PT-E4`, `PT-C4`, `PT-I4`),
               values_to = "PT4",
               names_to = "issue4") %>%
  mutate(issue4 = recode(issue4, 
                         `PT-C4` = "COVID-19",
                         `PT-D4` = "Defense",
                         `PT-E4` = "Education",
                         `PT-I4` = "Immigration"),
        check = if_else(issue == issue1 & issue == issue2 & issue == issue3 & issue == issue4, 1,0)) %>%
  filter(check == 1) %>%
  select(`Gender` = F1,
         `Age` = F2,
         `Education` = F9,
         `Region`  = F3,
         `Urbanness`  = F4,
         `Migration Background`  = F5,
         `Job`  = F7,
         `Income`  = F8,
         `Position: Immigration (%)`  = PreT1,
         `Position: Immigration`  = PreT2,
         `Position: Defense`  = PreT3,
         `Position: Education`  = PreT4,
         `Position: Attitude Women in Politics`  = PreT5,
         `Ideology`  = PreT6,
         `Political Efficacy` = PreT7, 
         `Position: COVID-19 Policy` = PreT8,
         `DV: Trait Evaluation` = PT1,
         `DV: Favorability` = PT2,
         `DV: Representation` = PT3,
         `DV: Career Prospects` = PT4) %>%
  ggcorrmat(
  type = "robust", # correlation method
  cor.vars = c(everything()), # a range of variables can be selected
  colors = c(fig_cols[1], fig_cols[4], fig_cols[8]),
  title = "Correlalogram for Variables under Study"
)
```