### 2021-05-06 report/draft.Rmd
add draft

---
title: "Bend or Break? Experimental Evidence of Voters' Punishment for Striking Political Compromises"
date: "`r format(Sys.time(), '%d %B, %Y - %X (%Z)')`"
output: 
  pdf_document:
    number_sections: true
editor_options: 
  chunk_output_type: console
bibliography: references.bib
urlcolor: blue
linkcolor: blue
abstract: "_ADD Abstract_"
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
---
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
library(tidyverse)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
library(here)
```

\newpage

# Introduction

# Acceptance of Political Compromises

# Data, Measurement & Method

```{r dv, fig.align = 'center', out.width = "75%", fig.cap = "Descriptive Overview Dependent Variable"}
include_graphics(here("report/figures", "Dependent Variable-1.png"))
```

```{r treatment, fig.align = 'center', out.width = "75%", fig.cap = "Treatment Distribution"}
include_graphics(here("report/figures", "Independent Variables-1.png"))
```

```{r balance, fig.align = 'center', out.width = "75%", fig.cap = "Balance Checks"}
include_graphics(here("report/figures", "balance-checks-1.png"))
```

```{r descriptive-table, echo = FALSE}
library(tidyverse)
load(here("data/intermediate/cleaned_data.RData"))
source(here("src/analysis/data-for-analyses.R"))
means <- d %>%
  summarise(F4 = round(mean(F4, na.rm = T),0),
            F5 = round(mean(F5, na.rm = T), 0),
            F6 = max(names(table(F6))),
            F7 = round(mean(F7, na.rm = T),0),
            F8 = round(mean(F8, na.rm = T),0),
            F9 = round(mean(F9, na.rm = T),0))

d <- d %>%
  mutate(F4 = replace_na(F4, means$F4),
         F5 = replace_na(F5, means$F5),
         F6 = replace_na(F6, means$F6),
         F7 = replace_na(F7, means$F7),
         missing_F8 = if_else(is.na(F8), 1, 0),
         F8 = replace_na(F8, means$F8),
         F9 = replace_na(F9, means$F9))
source(here("src/analysis/descriptive-information-overview.R"))
kbl(descr, booktabs =T, caption = "\\label{tab:descr}Descriptive Information",
    digits = 2) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center")
rm(descr, d, means)  
```

# Bend or Break?

```{r h1, fig.align = 'center', out.width = "75%", fig.cap = "H1"}
include_graphics(here("report/figures", "h1-1.png"))
```

```{r h2, fig.align = 'center', out.width = "75%", fig.cap = "H2"}
include_graphics(here("report/figures", "h2-1.png"))
```

```{r h3, fig.align = 'center', out.width = "75%", fig.cap = "H3"}
include_graphics(here("report/figures", "h3-1.png"))
```

```{r h3p, fig.align = 'center', out.width = "75%", fig.cap = "H13 Pooled"}
include_graphics(here("report/figures", "h3-pooled-1.png"))
```

# Discussion

\newpage
# References