### 2021-05-07 report/draft.Rmd
adapt figures

---
title: "Bend or Break? Experimental Evidence of Voters' Punishment for Striking Political Compromises"
date: "`r format(Sys.time(), '%d %B, %Y - %X (%Z)')`"
output: 
  pdf_document:
   number_sections: true
   includes:
      in_header: header.tex
editor_options: 
  chunk_output_type: console
bibliography: references.bib
urlcolor: blue
linkcolor: blue
abstract: "_ADD Abstract_"
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
---
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
options(kableExtra.latex.load_packages = FALSE)
source(here::here("src/lib/functions.R"))
```

\newpage

# Introduction

# Acceptance of Political Compromises

**Gender hypothesis** (_H1a_):  Women politicians are evaluated lower when striking a political compromise compared to men politicians.

**Gender hypothesis** (_H1b_):  Women politicians are evaluated lower when not striking a political compromise compared to men politicians.

**Migration hypothesis** (_H2_): Politicians with a migration background are evaluated lower when striking a political compromise compared to politicians without a migration background.

**Intersection hypothesis** (_H3a_): Women politician with a migration background are evaluated lower when striking a political compromise compared to men politicians without a migration background.

**Intersection hypothesis** (_H3b_): Women politician with a migration background are evaluated lower when not striking a political compromise compared to men politicians without a migration background.

# Data, Measurement & Method
The survey experiment is conducted in the Netherlands between the 21st of April 2021 and the 4th of May 2021.
**WHY NL**

## Data
Participants were told that they are taking part in a survey to get an overview of how Dutch people form their views on politics.
The sample, recruited through [KiesKompas](https://www.kieskompas.nl/en/),  consists of 2,004 participants of 18 years and older.[^1]
Because Kieskompas works with non-random opt-in respondents, I have measured many demographic and political background variables (described in [Section 3.2](#measurement) under _Covariates_).
Balance checks are conducted to demonstrate whether certain categories are over represented in a certain experimental group (see [Section 3.3](#method), Figure \ref{fig:balance}).
The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the _Vrije Universiteit Amsterdam_ (see the approval [here](https://github.com/MarikenvdVelden/compromise-punish/blob/main/docs/EthicalApproval.pdf)).
To ensure good quality of our data, two attention checks (discussed in more detail in [Section 3.3 of the Pre-Analysis Plan](https://github.com/MarikenvdVelden/bend-break-compromise/blob/main/docs/pap/compromise-punishment-pap.pdf)) are included.
All respondents included in the analyses passed both attention checks.

[^1]: The sample is calculated based on a power analysis presented in the [Pre-Analysis Plan](https://github.com/MarikenvdVelden/bend-break-compromise/blob/main/docs/pap/compromise-punishment-pap.pdf).

## Measurement

**_Dependent Variables_**
I rely on four measures reflecting different aspects of how people judge politicians.
The evaluation of politicians' character traits are statements deriving from @aaldering2016political.
The other three neasures to evaluate the politician of the treatment are translated from @bauer2017punished -- i.e. favorability, representation, and career perspectives.
General favorability is shown to be related to vote choice [@elis2010dynamics], perceptions of represenation are related to legislative compromise and cooperation [@doherty2013whom], and career perspective is included because critiques of women are typically manifested in their leadership skills [@eagly2007female].

```{r dv, fig.align = 'center', out.width = "90%", fig.cap = "\\label{fig:descr_dv}Descriptive Overview Dependent Variable"}
include_graphics(here("report/figures", "Dependent Variable-1.png"))
```

**_Experimental Conditions_**
The stimuli in the experiment are news messages in the same style as the Dutch news website [nu.nl](https://www.nu.nl/).
In these news messages, we manipulate: a) the _gender_ of the politician (male vs. female); b) the _migration background_ of the politician (based on a Arabic sounding name (Rachid(a) Amezian) or a native Dutch sounding name (Karel/Karin van der Kleijn)); and c) whether the politician _striked a compromise or not_.
This creates a full `2*2*2` factorial experiment with four rounds (see Table \ref{tab:treat}).
Every round, the news message covers a new issue.
The issues are: "defense" (increase cooperation of the defense units), "immigration" (immigration stop in western countries), "education" (expansion creative elementary school activities), "COVID-19" (emergency legislation to combat the COVID-19 pandemic).
For every round, the politicians' name, and thereby gender and immigration background, are manipulated in the text and in the picture, as well as the decision to strike a compromise.

```{r stimulus, topcaption=TRUE}
d <- tibble(` ` = c(1:8),
            `Name Politician` = c(rep("Rachid Amezian", 2), rep("Karel van der Kleijn", 2),
                                  rep("Rachida Amezian", 2), rep("Karin van der Kleijn", 2)),
            Gender = c(rep("Men",4), rep("Women", 4)),
            `Migration Background` = c(rep("Yes", 2), rep("No", 2),
                                  rep("Yes", 2), rep("No", 2)),
            Compromise = c("Yes", "No", "Yes", "No", "Yes", "No", "Yes", "No"))

kbl(d, booktabs =T, caption = "\\label{tab:treat}Experimental Conditions",
    align = c("l", "l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = ".5cm") %>%
  column_spec(2, width = "3.2cm") %>%
  column_spec(3, width = "2.5cm")%>%
  column_spec(4, width = "3.5cm")%>%
  column_spec(5, width = "2.5cm")

```

The text of the treatment is a Dutch translation of the stimulus material developed by @bauer2017punished.
The authors have pre-tested the text of the stimuli material several times [see the [Web Appendix](https://static-content.springer.com/esm/art%3A10.1007%2Fs11109-016-9356-6/MediaObjects/11109_2016_9356_MOESM1_ESM.docx) of @bauer2017punished].
The photo's of their stimuli material can unfortunately not be used, because a) @bauer2017punished only investigate the difference between men and women politicians, using solely photo''s of caucasians; and b) the American flag was present in their material.
Instead, I have used photo's from local politicians of the city of The Hague that are not know to the general public. 
The men politicians are in real life representatives of the [liberal-progressive party D66](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/boutkabout-a.-abid-fractievertegenwoordiger-d66.htm) (man with immigrant background) and of the [christian democratic party CDA](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/groot-de-d.p.-daniel-fractievertegenwoordiger-cda.htm) (man without immigrant background).
The women politicians are in real life representatives of a [right-wing populist party _Hart voor Den Haag/ Groep de Mos_](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/davituliani-n.-nino-raadslid-hart-voor-den-haaggroep-de-mos.htm) (woman with migration background) and of the [social democratic party PvdA](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/holman-j.-janneke-raadslid-pvda.htm) (woman without immigrant background).
Thereby we have a spread over the ideological different parties present in Dutch politics.
The names for the politicians without immigration background are _Karel_ (man) and _Karin_ (woman) _van der Kleijn_, which are typical native Dutch names.
For the politicians with an immigration background, we picked _Rachid_ (man) and _Rachida_ (woman) _Amezian_.
All stimulus material is developed by the [Network Institute of the _Vrije Universiteit Amsterdam_](https://networkinstitute.org/) and can be accessed [here](https://github.com/MarikenvdVelden/compromise-punish/tree/main/docs/pap/treatments).

**_Covariates_**

```{r descriptive-table, echo = FALSE}
load(here("data/intermediate/cleaned_data.RData"))
source(here("src/analysis/data-for-analyses.R"))
means <- d %>%
  summarise(F4 = round(mean(F4, na.rm = T),0),
            F5 = round(mean(F5, na.rm = T), 0),
            F6 = max(names(table(F6))),
            F7 = round(mean(F7, na.rm = T),0),
            F8 = round(mean(F8, na.rm = T),0),
            F9 = round(mean(F9, na.rm = T),0))

d <- d %>%
  mutate(F4 = replace_na(F4, means$F4),
         F5 = replace_na(F5, means$F5),
         F6 = replace_na(F6, means$F6),
         F7 = replace_na(F7, means$F7),
         missing_F8 = if_else(is.na(F8), 1, 0),
         F8 = replace_na(F8, means$F8),
         F9 = replace_na(F9, means$F9))
source(here("src/analysis/descriptive-information-overview.R"))
kbl(descr, booktabs =T, caption = "\\label{tab:descr}Descriptive Information",
    digits = 2) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center")
rm(descr, means)  
```

## Method

```{r balance, fig.align = 'center', out.width = "95%", fig.cap = "\\label{fig:balance}Balance Checks"}
include_graphics(here("report/figures", "balance-checks-1.png"))
```

# Bend or Break?

## Pre-Registered Hypotheses
\newpage 
\blandscape 
```{r results-hs, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results}Results"}
include_graphics(here("report/figures", "save-patchwork-results-1.png"))
```
\elandscape

## Exploration

# Discussion

\newpage
# References