### 2021-05-07 src/data-processing/prep_data.Rmd

---
title: 'Prepare Data'
author: 
output: 
  github_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center",
                      fig.path=('../../report/figures/'))
library(printr)
options(digits = 2)
library(rmarkdown)
```

# Scripts
- [Required Packages &amp; Reproducibility](#required-packages-&amp;-reproducibility)
- [Tidy Data](#tidy-data)
- [Save Data for Analysis](#save-data-for-analysis)
- [Check Missing Values](#check-missing-values)
- [Visualization of Data](#visualization-of-data)
  - [Dependent Variable](#dependent-variable)
  - [Treatment Conditions](#treatment-conditions)
  - [Correlations Matrix](#correlations-matrix)

## Required Packages &amp; Reproducibility
```{r "env", message=FALSE, warning=F}
rm(list=ls())
source(here::here("src/lib/functions.R"))
renv::snapshot()
```

## Tidy Data
This code chuck downloads the data from Qualtrics using the API and cleans the raw data.

```{r "get data from Qualtrics", results="hide"}
source(here("src/data-processing/tidy_data.R"))
source(here("src/data-processing/Urbaness.R"))
```

## Save Data for Analysis
```{r "save data",eval=FALSE}
save(d, file = here("data/intermediate/cleaned_data.RData"))
```

## Visualization of Data

### Dependent Variable
```{r "Dependent Variable", echo=F, fig.width=10, fig.height=8}
d %>%
  mutate(PT_D1 = round((PT_D1_1 + PT_D1_2 + PT_D1_3 + PT_D1_4 + PT_D1_5 + PT_D1_6)/6, 0),
         PT_E1 = round((PT_E1_1 + PT_E1_2 + PT_E1_3 + PT_E1_4 + PT_E1_5 + PT_E1_6)/6, 0),
         PT_C1 = round((PT_C1_1 + PT_C1_2 + PT_C1_3 + PT_C1_4 + PT_C1_5 + PT_C1_6)/6, 0),
         PT_I1 = round((PT_I1_1 + PT_I1_2 + PT_I1_3 + PT_I1_4 + PT_I1_5 + PT_I1_6)/6, 0),
         PreT5 = round((PreT5_1 + PreT5_2 + PreT5_3 + PreT5_4)/4,0),
         PreT7 = round((PreT7_1 + PreT7_2 + PreT7_3 + PreT7_4 + PreT7_5 + PreT7_6 + PreT7_7)/7,0)) %>%
  pivot_longer(cols = c(name_D, name_I, name_E, name_C),
               values_to = "name",
               names_to = "issue") %>%
  drop_na(name) %>%
  mutate(issue = recode(issue, 
                        name_C = "COVID",
                        name_D = "Defense",
                        name_E = "Education",
                        name_I = "Immigration")) %>%
  pivot_longer(cols = c(PT_D1, PT_E1, PT_C1, PT_I1),
               values_to = "PT1",
               names_to = "issue1") %>%
  drop_na(PT1) %>%
  mutate(issue1 = recode(issue1, 
                        PT_C1 = "COVID",
                        PT_D1 = "Defense",
                        PT_E1 = "Education",
                        PT_I1 = "Immigration")) %>%
  pivot_longer(cols = c(PT_D2, PT_E2, PT_C2, PT_I2),
               values_to = "PT2",
               names_to = "issue2") %>%
  mutate(issue2 = recode(issue2, 
                        PT_C2 = "COVID",
                        PT_D2 = "Defense",
                        PT_E2 = "Education",
                        PT_I2 = "Immigration")) %>%
  pivot_longer(cols = c(PT_D3, PT_E3, PT_C3, PT_I3),
               values_to = "PT3",
               names_to = "issue3") %>%
  mutate(issue3 = recode(issue3, 
                         PT_C3 = "COVID",
                         PT_D3 = "Defense",
                         PT_E3 = "Education",
                         PT_I3 = "Immigration")) %>%
  pivot_longer(cols = c(PT_D4, PT_E4, PT_C4, PT_I4),
               values_to = "PT4",
               names_to = "issue4") %>%
  mutate(issue4 = recode(issue4, 
                         PT_C4 = "COVID",
                         PT_D4 = "Defense",
                         PT_E4 = "Education",
                         PT_I4 = "Immigration"),
        check = if_else(issue == issue1 & issue == issue2 & issue == issue3 & issue == issue4, 1,0)) %>%
  filter(check == 1) %>%
  select(PT1, PT2, PT3, PT4, issue = issue1) %>%
  mutate(PT1 = PT1 * 2) %>%
  #mutate_at(c("PT1", "PT2", "PT3", "PT4"), ~(scale(.) %>% as.vector)) %>%
  pivot_longer(cols = PT1:PT4,
               names_to = "dvs") %>%
  mutate(dvs = recode(dvs,
                       PT1 = "DV: Trait Evaluation",
                       PT2 = "DV: Favorability",
                       PT3 = "DV: Representation",
                       PT4 = "DV: Career Prospects")) %>%
  group_by(issue, dvs) %>%
  summarise(means = mean(value, na.rm=T),
            stdev = sd(value, na.rm=T)) %>%
  mutate(lower = means - (1.56 * stdev),
         upper = means + (1.56 * stdev)) %>%
  ggplot(aes(x = dvs, y = means, colour = issue,
             ymin = lower, ymax = upper, label = issue)) +
  geom_point(position = position_dodge(.5)) + 
  geom_errorbar(position = position_dodge(.5), width = 0) +
  geom_text(aes(y = upper +.5), 
            position = position_dodge(.5),na.rm = TRUE) + 
  labs(y = "0 (negative) - 10 (positive)", 
       x = "") +
  geom_hline(yintercept = 5, linetype = "dotted", color = "darkgrey") +
  theme_minimal() +
  scale_color_manual(values = fig_cols) +
  coord_flip() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="none",
        legend.title = element_blank()) 
```

### Treatment Conditions
```{r "Independent Variables", echo=F, message=FALSE, warning=FALSE,fig.width=10, fig.height=10}

d %>%
  mutate(PT_D1 = round((PT_D1_1 + PT_D1_2 + PT_D1_3 + PT_D1_4 + PT_D1_5 + PT_D1_6)/6, 0),
         PT_E1 = round((PT_E1_1 + PT_E1_2 + PT_E1_3 + PT_E1_4 + PT_E1_5 + PT_E1_6)/6, 0),
         PT_C1 = round((PT_C1_1 + PT_C1_2 + PT_C1_3 + PT_C1_4 + PT_C1_5 + PT_C1_6)/6, 0),
         PT_I1 = round((PT_I1_1 + PT_I1_2 + PT_I1_3 + PT_I1_4 + PT_I1_5 + PT_I1_6)/6, 0),
         PreT5 = round((PreT5_1 + PreT5_2 + PreT5_3 + PreT5_4)/4,0),
         PreT7 = round((PreT7_1 + PreT7_2 + PreT7_3 + PreT7_4 + PreT7_5 + PreT7_6 + PreT7_7)/7,0)) %>%
  pivot_longer(cols = c(name_D, name_I, name_E, name_C),
               values_to = "name",
               names_to = "issue") %>%
  drop_na(name) %>%
  mutate(issue = recode(issue, 
                        name_C = "COVID_19",
                        name_D = "Defense",
                        name_E = "Education",
                        name_I = "Immigration")) %>%
  pivot_longer(cols = c(PT_D1, PT_E1, PT_C1, PT_I1),
               values_to = "PT1",
               names_to = "issue1") %>%
  drop_na(PT1) %>%
  mutate(issue1 = recode(issue1, 
                        PT_C1 = "COVID_19",
                        PT_D1 = "Defense",
                        PT_E1 = "Education",
                        PT_I1 = "Immigration")) %>%
  pivot_longer(cols = c(PT_D2, PT_E2, PT_C2, PT_I2),
               values_to = "PT2",
               names_to = "issue2") %>%
  mutate(issue2 = recode(issue2, 
                        PT_C2 = "COVID_19",
                        PT_D2 = "Defense",
                        PT_E2 = "Education",
                        PT_I2 = "Immigration")) %>%
  pivot_longer(cols = c(PT_D3, PT_E3, PT_C3, PT_I3),
               values_to = "PT3",
               names_to = "issue3") %>%
  mutate(issue3 = recode(issue3, 
                         PT_C3 = "COVID_19",
                         PT_D3 = "Defense",
                         PT_E3 = "Education",
                         PT_I3 = "Immigration")) %>%
  pivot_longer(cols = c(PT_D4, PT_E4, PT_C4, PT_I4),
               values_to = "PT4",
               names_to = "issue4") %>%
  mutate(issue4 = recode(issue4, 
                         PT_C4 = "COVID_19",
                         PT_D4 = "Defense",
                         PT_E4 = "Education",
                         PT_I4 = "Immigration"),
        check = if_else(issue == issue1 & issue == issue2 & issue == issue3 & issue == issue4, 1,0)) %>%
  filter(check == 1) %>%
  mutate(gender = recode(name,
                         `Karel van der Kleijn` = 0,
                         `Karin van der Kleijn` = 1,
                         `Rachid Amezian` = 0,
                         `Rachida Amezian` = 1),
         immigration = recode(name,
                              `Karel van der Kleijn` = 0,
                              `Karin van der Kleijn` = 0,
                              `Rachid Amezian` = 1,
                              `Rachida Amezian` = 1),
         intersection = recode(name,
                               `Karel van der Kleijn` = 0,
                               `Karin van der Kleijn` = 0,
                               `Rachid Amezian` = 0,
                               `Rachida Amezian` = 1),
         compromise = if_else(compromis == "geen compromis", "Compromise", "No Compromise")) %>%
  unite("treatment", c("name", "compromise"), sep = "\n ") %>%
  select(treatment, issue) %>%
  group_by(issue, treatment) %>%
  summarise(n = n()) %>%
  mutate(perc = round(n/sum(n),2),
         perc2 = paste0(perc*100, "%")) %>%
  ggplot(aes(x = perc, y = issue, fill = treatment, label = perc2)) +
  geom_col(position = 'fill', width=0.75) + 
  geom_text(position = position_fill(vjust=0.5), color = "white", fontface = "bold") +
  labs(y = "", x = "", title = "Distribution of Treatments") +
  scale_x_continuous(labels=scales::percent) +
  theme_minimal() +
  scale_fill_manual(values = fig_cols) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position="bottom",
        legend.title = element_blank()) +
  guides(fill=guide_legend(ncol=4,byrow=TRUE))
```

### Correlations Matrix
```{r "Correlations Matrix", echo=F, message=T, warning=T, fig.width=15, fig.height=15}
d %>%
  mutate(PT_D1 = round((PT_D1_1 + PT_D1_2 + PT_D1_3 + PT_D1_4 + PT_D1_5 + PT_D1_6)/6, 0),
         PT_E1 = round((PT_E1_1 + PT_E1_2 + PT_E1_3 + PT_E1_4 + PT_E1_5 + PT_E1_6)/6, 0),
         PT_C1 = round((PT_C1_1 + PT_C1_2 + PT_C1_3 + PT_C1_4 + PT_C1_5 + PT_C1_6)/6, 0),
         PT_I1 = round((PT_I1_1 + PT_I1_2 + PT_I1_3 + PT_I1_4 + PT_I1_5 + PT_I1_6)/6, 0),
         PreT5 = round((PreT5_1 + PreT5_2 + PreT5_3 + PreT5_4)/4,0),
         PreT7 = round((PreT7_1 + PreT7_2 + PreT7_3 + PreT7_4 + PreT7_5 + PreT7_6 + PreT7_7)/7,0)) %>%
  pivot_longer(cols = c(name_D, name_I, name_E, name_C),
               values_to = "name",
               names_to = "issue") %>%
  drop_na(name) %>%
  mutate(issue = recode(issue, 
                        `name_C` = "COVID_19",
                        `name_D` = "Defense",
                        `name_E` = "Education",
                        `name_I` = "Immigration")) %>%
  pivot_longer(cols = c(`PT_D1`, `PT_E1`, `PT_C1`, `PT_I1`),
               values_to = "PT1",
               names_to = "issue1") %>%
  drop_na(PT1) %>%
  mutate(issue1 = recode(issue1, 
                        `PT_C1` = "COVID_19",
                        `PT_D1` = "Defense",
                        `PT_E1` = "Education",
                        `PT_I1` = "Immigration")) %>%
  pivot_longer(cols = c(`PT_D2`, `PT_E2`, `PT_C2`, `PT_I2`),
               values_to = "PT2",
               names_to = "issue2") %>%
  mutate(issue2 = recode(issue2, 
                        `PT_C2` = "COVID_19",
                        `PT_D2` = "Defense",
                        `PT_E2` = "Education",
                        `PT_I2` = "Immigration")) %>%
  pivot_longer(cols = c(`PT_D3`, `PT_E3`, `PT_C3`, `PT_I3`),
               values_to = "PT3",
               names_to = "issue3") %>%
  mutate(issue3 = recode(issue3, 
                         `PT_C3` = "COVID_19",
                         `PT_D3` = "Defense",
                         `PT_E3` = "Education",
                         `PT_I3` = "Immigration")) %>%
  pivot_longer(cols = c(`PT_D4`, `PT_E4`, `PT_C4`, `PT_I4`),
               values_to = "PT4",
               names_to = "issue4") %>%
  mutate(issue4 = recode(issue4, 
                         `PT_C4` = "COVID_19",
                         `PT_D4` = "Defense",
                         `PT_E4` = "Education",
                         `PT_I4` = "Immigration"),
        check = if_else(issue == issue1 & issue == issue2 & issue == issue3 & issue == issue4, 1,0)) %>%
  filter(check == 1) %>%
  select(`Gender` = F1,
         `Age` = F2,
         `Education` = F9,
         `Region`  = F3,
         `Urbanness`  = F4,
         `Migration Background`  = F5,
         `Job`  = F7,
         `Income`  = F8,
         `Position: Immigration (%)`  = PreT1,
         `Position: Immigration`  = PreT2,
         `Position: Defense`  = PreT3,
         `Position: Education`  = PreT4,
         `Position: Attitude Women in Politics`  = PreT5,
         `Ideology`  = PreT6,
         `Political Efficacy` = PreT7, 
         `Position: COVID-19 Policy` = PreT8,
         `DV: Trait Evaluation` = PT1,
         `DV: Favorability` = PT2,
         `DV: Representation` = PT3,
         `DV: Career Prospects` = PT4) %>%
  ggcorrmat(
  type = "robust", # correlation method
  cor.vars = c(everything()), # a range of variables can be selected
  colors = c(fig_cols[1], fig_cols[4], fig_cols[8]),
  title = "Correlalogram for Variables under Study"
)
```