### 2021-05-07 report/draft.Rmd
add exploratory results

---
title: "Bend or Break? Experimental Evidence of Voters' Punishment for Striking Political Compromises"
date: "`r format(Sys.time(), '%d %B, %Y - %X (%Z)')`"
output: 
  pdf_document:
   number_sections: true
   includes:
      in_header: header.tex
editor_options: 
  chunk_output_type: console
bibliography: references.bib
urlcolor: blue
linkcolor: blue
abstract: "_ADD Abstract_"
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
---
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
options(kableExtra.latex.load_packages = FALSE)
source(here::here("src/lib/functions.R"))
```

\newpage

# Introduction

# Acceptance of Political Compromises

**Gender hypothesis** (_H1a_):  Women politicians are evaluated lower when striking a political compromise compared to men politicians.

**Gender hypothesis** (_H1b_):  Women politicians are evaluated lower when not striking a political compromise compared to men politicians.

**Migration hypothesis** (_H2_): Politicians with a migration background are evaluated lower when striking a political compromise compared to politicians without a migration background.

**Intersection hypothesis** (_H3a_): Women politician with a migration background are evaluated lower when striking a political compromise compared to men politicians without a migration background.

**Intersection hypothesis** (_H3b_): Women politician with a migration background are evaluated lower when not striking a political compromise compared to men politicians without a migration background.

# Data, Measurement & Method
The survey experiment is conducted in the Netherlands between the 21st of April 2021 and the 4th of May 2021.
**WHY NL**

## Data
Participants were told that they are taking part in a survey to get an overview of how Dutch people form their views on politics.
The sample, recruited through [KiesKompas](https://www.kieskompas.nl/en/),  consists of 2,004 participants of 18 years and older.[^1]
Because Kieskompas works with non-random opt-in respondents, I have measured many demographic and political background variables (described in [Section 3.2](#measurement) under _Covariates_).
Balance checks are conducted to demonstrate whether certain categories are over represented in a certain experimental group (see [Section 3.3](#method), Figure \ref{fig:balance}).
The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the _Vrije Universiteit Amsterdam_ (see the approval [here](https://github.com/MarikenvdVelden/compromise-punish/blob/main/docs/EthicalApproval.pdf)).
To ensure good quality of our data, two attention checks (discussed in more detail in [Section 3.3 of the Pre-Analysis Plan](https://github.com/MarikenvdVelden/bend-break-compromise/blob/main/docs/pap/compromise-punishment-pap.pdf)) are included.
All respondents included in the analyses passed both attention checks.

[^1]: The sample is calculated based on a power analysis presented in the [Pre-Analysis Plan](https://github.com/MarikenvdVelden/bend-break-compromise/blob/main/docs/pap/compromise-punishment-pap.pdf).

## Measurement

**_Dependent Variables_**
I rely on four measures reflecting different aspects of how people judge politicians.
The evaluation of politicians' character traits are statements deriving from @aaldering2016political.
The other three neasures to evaluate the politician of the treatment are translated from @bauer2017punished -- i.e. favorability, representation, and career perspectives.
General favorability is shown to be related to vote choice [@elis2010dynamics], perceptions of represenation are related to legislative compromise and cooperation [@doherty2013whom], and career perspective is included because critiques of women are typically manifested in their leadership skills [@eagly2007female].

**_Experimental Conditions_**
The stimuli in the experiment are news messages in the same style as the Dutch news website [nu.nl](https://www.nu.nl/).
In these news messages, we manipulate: a) the _gender_ of the politician (male vs. female); b) the _migration background_ of the politician (based on a Arabic sounding name (Rachid(a) Amezian) or a native Dutch sounding name (Karel/Karin van der Kleijn)); and c) whether the politician _striked a compromise or not_.
This creates a full `2*2*2` factorial experiment with four rounds (see Table \ref{tab:treat}).
Every round, the news message covers a new issue.
The issues are: "defense" (increase cooperation of the defense units), "immigration" (immigration stop in western countries), "education" (expansion creative elementary school activities), "COVID-19" (emergency legislation to combat the COVID-19 pandemic).
For every round, the politicians' name, and thereby gender and immigration background, are manipulated in the text and in the picture, as well as the decision to strike a compromise.

```{r stimulus, topcaption=TRUE}
d <- tibble(` ` = c(1:8),
            `Name Politician` = c(rep("Rachid Amezian", 2), rep("Karel van der Kleijn", 2),
                                  rep("Rachida Amezian", 2), rep("Karin van der Kleijn", 2)),
            Gender = c(rep("Men",4), rep("Women", 4)),
            `Migration Background` = c(rep("Yes", 2), rep("No", 2),
                                  rep("Yes", 2), rep("No", 2)),
            Compromise = c("Yes", "No", "Yes", "No", "Yes", "No", "Yes", "No"))

kbl(d, booktabs =T, caption = "\\label{tab:treat}Experimental Conditions",
    align = c("l", "l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = ".5cm") %>%
  column_spec(2, width = "3.2cm") %>%
  column_spec(3, width = "2.5cm")%>%
  column_spec(4, width = "3.5cm")%>%
  column_spec(5, width = "2.5cm")

```

The text of the treatment is a Dutch translation of the stimulus material developed by @bauer2017punished.
The authors have pre-tested the text of the stimuli material several times [see the [Web Appendix](https://static-content.springer.com/esm/art%3A10.1007%2Fs11109-016-9356-6/MediaObjects/11109_2016_9356_MOESM1_ESM.docx) of @bauer2017punished].
The photo's of their stimuli material can unfortunately not be used, because a) @bauer2017punished only investigate the difference between men and women politicians, using solely photo''s of caucasians; and b) the American flag was present in their material.
Instead, I have used photo's from local politicians of the city of The Hague that are not know to the general public. 
The men politicians are in real life representatives of the [liberal-progressive party D66](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/boutkabout-a.-abid-fractievertegenwoordiger-d66.htm) (man with immigrant background) and of the [christian democratic party CDA](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/groot-de-d.p.-daniel-fractievertegenwoordiger-cda.htm) (man without immigrant background).
The women politicians are in real life representatives of a [right-wing populist party _Hart voor Den Haag/ Groep de Mos_](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/davituliani-n.-nino-raadslid-hart-voor-den-haaggroep-de-mos.htm) (woman with migration background) and of the [social democratic party PvdA](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/holman-j.-janneke-raadslid-pvda.htm) (woman without immigrant background).
Thereby we have a spread over the ideological different parties present in Dutch politics.
The names for the politicians without immigration background are _Karel_ (man) and _Karin_ (woman) _van der Kleijn_, which are typical native Dutch names.
For the politicians with an immigration background, we picked _Rachid_ (man) and _Rachida_ (woman) _Amezian_.
All stimulus material is developed by the [Network Institute of the _Vrije Universiteit Amsterdam_](https://networkinstitute.org/) and can be accessed [here](https://github.com/MarikenvdVelden/compromise-punish/tree/main/docs/pap/treatments).

**_Covariates_**

```{r descriptive-table, echo = FALSE}
load(here("data/intermediate/cleaned_data.RData"))
source(here("src/analysis/data-for-analyses.R"))
means <- d %>%
  summarise(F4 = round(mean(F4, na.rm = T),0),
            F5 = round(mean(F5, na.rm = T), 0),
            F6 = max(names(table(F6))),
            F7 = round(mean(F7, na.rm = T),0),
            F8 = round(mean(F8, na.rm = T),0),
            F9 = round(mean(F9, na.rm = T),0))

d <- d %>%
  mutate(F4 = replace_na(F4, means$F4),
         F5 = replace_na(F5, means$F5),
         F6 = replace_na(F6, means$F6),
         F7 = replace_na(F7, means$F7),
         missing_F8 = if_else(is.na(F8), 1, 0),
         F8 = replace_na(F8, means$F8),
         F9 = replace_na(F9, means$F9))
source(here("src/analysis/descriptive-information-overview.R"))
kbl(descr, booktabs =T, caption = "\\label{tab:descr}Descriptive Information",
    digits = 2) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center")
rm(descr, means)  
```

## Method

```{r balance, fig.align = 'center', out.width = "95%", fig.cap = "\\label{fig:balance}Balance Checks"}
include_graphics(here("report/figures", "balance-checks-1.png"))
```

# Bend or Break?

## Pre-Registered Hypotheses

**_Gender Hypothesis_**
H1a states that women politicians are evaluated lower when striking a political compromise compared to men politicians. 
I inspect this using four different measures of how people can judge the quality of a political representative. 
The upper-left panel of Figure \ref{fig:results} demonstrates that people's evaluations of women politicians compared to men politicians differ between measures -- displayed on the y-axis of the figure -- and between issue areas -- displayed with different colors in the figure.

- Looking at trait evaluation, only women who compromise on the issue migration are evaluated more negatively then men, yet the coefficient is not statistically significant. For the other three issues -- COVID, defense, and education -- the coefficients are positive, yet not statistically significant.
- The same pattern is shown for how well people feel represented by the politician. Women, compared to men, are lower evaluated for the issue migration, but higher evaluated for the other three issues. Only the effect of education is statistically significant. This indicates that if female polticians compromise and vote in favor of the bill on extending extra-curricular activities at elementary schools, people feel better represented by this politician compared to if the same compromise was made by a male politician.
- Regarding feelings of favorability towards the politician, there is no statistical significant difference between male or female politicians who strike a compromise. The same holds for the politicians' career prospects.

H1b states that women politicians are evaluated lower when not striking a political compromise compared to men politicians. 
The upper-center-left panel of Figure \ref{fig:results} demonstrates that almost all coefficients are positive -- i.e. going against the hypothesized directions.

- When women politicians stand their grounds, they are evaluated higher on their leadership skills than their male counterparts. This finding holds accross issues, but is not statistically significant for the issue immigration.
- In addition, compared to male politicians, female politicians that stand their ground make people feel better represented, except when this is about an education-related policy. This finding is statistically significant for the issues immigration and defense. 
- Moreover, female politicians are more favourable when standing their grounds. This finding is only statistically significant when women, compared to men, vote against a bill on intensifying cooperation of the army services.
- Also, not striking a compromise is considered good for a women's career prospects, yet these findings are not statistically significant.

**_Migration Hypothesis_**
This hypothesis states that politicians with a migration background are evaluated lower when striking a political compromise compared to politicians without a migration background. 
The upper-center-right panel of Figure \ref{fig:results} demonstrates that politicians with a migration background compared to politicians without a migration background are more positively evaluated on some measures, but more negatively evaluated on other measures.

- Looking at their leadership skills (trait evaluation), we see that for the issues COVID-19, defense, and education, there is no difference between politicians with or without a migration background. However, for the issue of migration, we see that if politicians with a migration background compromise -- i.e. voting in favor of the bill proposing to put a stop to migration to Western countries -- they are considered better political leaders compared to politicians without a migration background.
- Regarding how well people feel represented, there is no difference between politicians with and without a migration background.
- Looking at how favorable people feel towards a politcian, we see that people feel more favorable towards politician with a migration background when they compromise on issues related to immigration and education compared to politicians without a migration background, but not for issues related to COVID-19 and defense.
- For career perspectives, people rate the career prospects more negative when a politician with a migration background compromises on the issue migration. For the other issues, there is no difference between politicians with and without a migration background.

\newpage 
\blandscape 
```{r results-main, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results}Results"}
include_graphics(here("report/figures", "save-patchwork-results-1.png"))
```
\elandscape

**_Intersection Hypothesis_** Hypothesis 3a states that women politician with a migration background are evaluated lower when striking a political compromise compared to men politicians without a migration background.
This hypothesis is tested in the lower-left panel of Figure \ref{fig:results} for the separate issues, and in the lower-center-right panel of Figure \ref{fig:results} pooling the issues.

- Trait: no differences for COVID-19 and immigration, insignificant positive effects for defense and education, pooling significant positive effect
- Representation: Positive and signicant for education, nothing for the other issues, pooling significant positive effect
- Favorability:  Positive and signicant for education, nothing for the other issues, pooling significant positive effect
- Career prospects: negative and significant for immigration, nothing for the other issues or in the pooled analysis.

Hypothesis 3b states that women politician with a migration background are evaluated lower when not striking a political compromise compared to men politicians without a migration background.
This hypothesis is tested in the lower-center-left panel of Figure \ref{fig:results} for the separate issues, and in the lower-right panel of Figure \ref{fig:results} pooling the issues.

- Trait: positive and significant for COVID-19, nothing for the other issues or in the pooled analysis.
- Representation: no differences
- Favorability: positive and significant for COVID-19, nothing for the other issues or in the pooled analysis.
- Career prospects:negative and significant for defense,  nothing for the other issues or in the pooled analysis.

Paragraph on what it means for theory, and justification for explorations (below).

## Exploration

**_Ideology_**
\newpage 
\blandscape 
```{r results-ideology, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-ideology}Results: Interaction with Ideology"}
include_graphics(here("report/figures", "explore-ideology-1.png"))
```
\elandscape

**_Willingness to Accept Compromises_**
\newpage 
\blandscape 
```{r results-wtac, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-wtac}Results: Interaction with Willingness to Accept Compromise"}
include_graphics(here("report/figures", "explore-wtac-1.png"))
```
\elandscape

**_Attitude towards Women in Politics_**
\newpage 
\blandscape 
```{r results-women-politics, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-women-politics}Results: Interaction with Attitudes towards Women in Politics"}
include_graphics(here("report/figures", "explore-women-politics-1.png"))
```
\elandscape

**_Position on Political Issues_**
\newpage 
\blandscape 
```{r results-covid, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-covid}Results: Interaction with Position on COVID-19"}
include_graphics(here("report/figures", "explore-covid-1.png"))
```
\elandscape

\newpage 
\blandscape 
```{r results-defense, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-defense}Results: Interaction with Position on Defense"}
include_graphics(here("report/figures", "explore-defense-1.png"))
```
\elandscape

\newpage 
\blandscape 
```{r results-education, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-education}Results: Interaction with Position on Education"}
include_graphics(here("report/figures", "explore-education-1.png"))
```
\elandscape

\newpage 
\blandscape 
```{r results-immigration, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-immigration}Results: Interaction with Position on Immigration"}
include_graphics(here("report/figures", "explore-immigration-1.png"))
```
\elandscape

# Discussion

\newpage
# References