### 2021-05-11 report/draft.Rmd
add interpretations

---
title: "Bend or Break? Experimental Evidence of Voters' Punishment for Striking Political Compromises"
date: "`r format(Sys.time(), '%d %B, %Y - %X (%Z)')`"
output: 
  pdf_document:
   number_sections: true
   includes:
      in_header: header.tex
editor_options: 
  chunk_output_type: console
bibliography: references.bib
urlcolor: blue
linkcolor: blue
abstract: "_ADD Abstract_"
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
---
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
options(kableExtra.latex.load_packages = FALSE)
source(here::here("src/lib/functions.R"))
```

\newpage

# Introduction

# Acceptance of Political Compromises

**Gender hypothesis** (_H1a_):  Women politicians are evaluated lower when striking a political compromise compared to men politicians.

**Gender hypothesis** (_H1b_):  Women politicians are evaluated lower when not striking a political compromise compared to men politicians.

**Migration hypothesis** (_H2_): Politicians with a migration background are evaluated lower when striking a political compromise compared to politicians without a migration background.

**Intersection hypothesis** (_H3a_): Women politician with a migration background are evaluated lower when striking a political compromise compared to men politicians without a migration background.

**Intersection hypothesis** (_H3b_): Women politician with a migration background are evaluated lower when not striking a political compromise compared to men politicians without a migration background.

# Data, Measurement & Method
The survey experiment is conducted in the Netherlands between the 21st of April 2021 and the 4th of May 2021.
**WHY NL**

## Data
Participants were told that they are taking part in a survey to get an overview of how Dutch people form their views on politics.
The sample, recruited through [KiesKompas](https://www.kieskompas.nl/en/),  consists of 2,004 participants of 18 years and older.[^1]
Because Kieskompas works with non-random opt-in respondents, I have measured many demographic and political background variables (described in [Section 3.2](#measurement) under _Covariates_).
Balance checks are conducted to demonstrate whether certain categories are over represented in a certain experimental group (see [Section 3.3](#method), Figure \ref{fig:balance}).
The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the _Vrije Universiteit Amsterdam_ (see the approval [here](https://github.com/MarikenvdVelden/compromise-punish/blob/main/docs/EthicalApproval.pdf)).
To ensure good quality of our data, two attention checks (discussed in more detail in [Section 3.3 of the Pre-Analysis Plan](https://github.com/MarikenvdVelden/bend-break-compromise/blob/main/docs/pap/compromise-punishment-pap.pdf)) are included.
All respondents included in the analyses passed both attention checks.

[^1]: The sample is calculated based on a power analysis presented in the [Pre-Analysis Plan](https://github.com/MarikenvdVelden/bend-break-compromise/blob/main/docs/pap/compromise-punishment-pap.pdf).

## Measurement

**_Dependent Variables_**
I rely on four measures reflecting different aspects of how people judge politicians.
The evaluation of politicians' character traits are statements deriving from @aaldering2016political.
The other three neasures to evaluate the politician of the treatment are translated from @bauer2017punished -- i.e. favorability, representation, and career perspectives.
General favorability is shown to be related to vote choice [@elis2010dynamics], perceptions of represenation are related to legislative compromise and cooperation [@doherty2013whom], and career perspective is included because critiques of women are typically manifested in their leadership skills [@eagly2007female].

**_Experimental Conditions_**
The stimuli in the experiment are news messages in the same style as the Dutch news website [nu.nl](https://www.nu.nl/).
In these news messages, we manipulate: a) the _gender_ of the politician (male vs. female); b) the _migration background_ of the politician (based on a Arabic sounding name (Rachid(a) Amezian) or a native Dutch sounding name (Karel/Karin van der Kleijn)); and c) whether the politician _striked a compromise or not_.
This creates a full `2*2*2` factorial experiment with four rounds (see Table \ref{tab:treat}).
Every round, the news message covers a new issue.
The issues are: "defense" (increase cooperation of the defense units), "immigration" (immigration stop in western countries), "education" (expansion creative elementary school activities), "COVID-19" (emergency legislation to combat the COVID-19 pandemic).
For every round, the politicians' name, and thereby gender and immigration background, are manipulated in the text and in the picture, as well as the decision to strike a compromise.

```{r stimulus, topcaption=TRUE}
d <- tibble(` ` = c(1:8),
            `Name Politician` = c(rep("Rachid Amezian", 2), rep("Karel van der Kleijn", 2),
                                  rep("Rachida Amezian", 2), rep("Karin van der Kleijn", 2)),
            Gender = c(rep("Men",4), rep("Women", 4)),
            `Migration Background` = c(rep("Yes", 2), rep("No", 2),
                                  rep("Yes", 2), rep("No", 2)),
            Compromise = c("Yes", "No", "Yes", "No", "Yes", "No", "Yes", "No"))

kbl(d, booktabs =T, caption = "\\label{tab:treat}Experimental Conditions",
    align = c("l", "l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = ".5cm") %>%
  column_spec(2, width = "3.2cm") %>%
  column_spec(3, width = "2.5cm")%>%
  column_spec(4, width = "3.5cm")%>%
  column_spec(5, width = "2.5cm")

```

The text of the treatment is a Dutch translation of the stimulus material developed by @bauer2017punished.
The authors have pre-tested the text of the stimuli material several times [see the [Web Appendix](https://static-content.springer.com/esm/art%3A10.1007%2Fs11109-016-9356-6/MediaObjects/11109_2016_9356_MOESM1_ESM.docx) of @bauer2017punished].
The photo's of their stimuli material can unfortunately not be used, because a) @bauer2017punished only investigate the difference between men and women politicians, using solely photo''s of caucasians; and b) the American flag was present in their material.
Instead, I have used photo's from local politicians of the city of The Hague that are not know to the general public. 
The men politicians are in real life representatives of the [liberal-progressive party D66](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/boutkabout-a.-abid-fractievertegenwoordiger-d66.htm) (man with immigrant background) and of the [christian democratic party CDA](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/groot-de-d.p.-daniel-fractievertegenwoordiger-cda.htm) (man without immigrant background).
The women politicians are in real life representatives of a [right-wing populist party _Hart voor Den Haag/ Groep de Mos_](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/davituliani-n.-nino-raadslid-hart-voor-den-haaggroep-de-mos.htm) (woman with migration background) and of the [social democratic party PvdA](https://www.denhaag.nl/nl/bestuur-en-organisatie/gemeenteraad/raadsleden-en-fractievertegenwoordigers/holman-j.-janneke-raadslid-pvda.htm) (woman without immigrant background).
Thereby we have a spread over the ideological different parties present in Dutch politics.
The names for the politicians without immigration background are _Karel_ (man) and _Karin_ (woman) _van der Kleijn_, which are typical native Dutch names.
For the politicians with an immigration background, we picked _Rachid_ (man) and _Rachida_ (woman) _Amezian_.
All stimulus material is developed by the [Network Institute of the _Vrije Universiteit Amsterdam_](https://networkinstitute.org/) and can be accessed [here](https://github.com/MarikenvdVelden/compromise-punish/tree/main/docs/pap/treatments).

**_Covariates_**

```{r descriptive-table, echo = FALSE}
load(here("data/intermediate/cleaned_data.RData"))
source(here("src/analysis/data-for-analyses.R"))
means <- d %>%
  summarise(F4 = round(mean(F4, na.rm = T),0),
            F5 = round(mean(F5, na.rm = T), 0),
            F6 = max(names(table(F6))),
            F7 = round(mean(F7, na.rm = T),0),
            F8 = round(mean(F8, na.rm = T),0),
            F9 = round(mean(F9, na.rm = T),0))

d <- d %>%
  mutate(F4 = replace_na(F4, means$F4),
         F5 = replace_na(F5, means$F5),
         F6 = replace_na(F6, means$F6),
         F7 = replace_na(F7, means$F7),
         missing_F8 = if_else(is.na(F8), 1, 0),
         F8 = replace_na(F8, means$F8),
         F9 = replace_na(F9, means$F9))
source(here("src/analysis/descriptive-information-overview.R"))
kbl(descr, booktabs =T, caption = "\\label{tab:descr}Descriptive Information",
    digits = 2) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center")  %>%
   column_spec(1, width = "5.5cm") %>%
   column_spec(2, width = "2cm") %>%
   column_spec(3, width = "1.5cm") %>%
   column_spec(4, width = "2cm") %>%
   column_spec(5, width = "2cm")
rm(descr, means)  
```

## Method

```{r balance, fig.align = 'center', out.width = "95%", fig.cap = "\\label{fig:balance}Balance Checks"}
include_graphics(here("report/figures", "balance-checks-1.png"))
```

# Bend or Break?

## Pre-Registered Hypotheses

**_Gender Hypothesis_**
H1a states that women politicians are evaluated lower when striking a political compromise compared to men politicians. 
I inspect this using four different measures of how people can judge the quality of a political representative. 
The upper-left panel of Figure \ref{fig:results} demonstrates that people's evaluations of women politicians compared to men politicians differ between measures -- displayed on the y-axis of the figure -- and between issue areas -- displayed with different colors in the figure.

- Looking at trait evaluation, only women who compromise on the issue migration are evaluated more negatively then men, yet the coefficient is not statistically significant. For the other three issues -- COVID, defense, and education -- the coefficients are positive, yet not statistically significant.
- The same pattern is shown for how well people feel represented by the politician. Women, compared to men, are lower evaluated for the issue migration, but higher evaluated for the other three issues. Only the effect of education is statistically significant. This indicates that if female polticians compromise and vote in favor of the bill on extending extra-curricular activities at elementary schools, people feel better represented by this politician compared to if the same compromise was made by a male politician.
- Regarding feelings of favorability towards the politician, there is no statistical significant difference between male or female politicians who strike a compromise. The same holds for the politicians' career prospects.

H1b states that women politicians are evaluated lower when not striking a political compromise compared to men politicians. 
The upper-center-left panel of Figure \ref{fig:results} demonstrates that almost all coefficients are positive -- i.e. going against the hypothesized directions.

- When women politicians stand their grounds, they are evaluated higher on their leadership skills than their male counterparts. This finding holds accross issues, but is not statistically significant for the issue immigration.
- In addition, compared to male politicians, female politicians that stand their ground make people feel better represented, except when this is about an education-related policy. This finding is statistically significant for the issues immigration and defense. 
- Moreover, female politicians are more favourable when standing their grounds. This finding is only statistically significant when women, compared to men, vote against a bill on intensifying cooperation of the army services.
- Also, not striking a compromise is considered good for a women's career prospects, yet these findings are not statistically significant.

**_Migration Hypothesis_**
This hypothesis states that politicians with a migration background are evaluated lower when striking a political compromise compared to politicians without a migration background. 
The upper-center-right panel of Figure \ref{fig:results} demonstrates that politicians with a migration background compared to politicians without a migration background are more positively evaluated on some measures, but more negatively evaluated on other measures.

- Looking at their leadership skills (trait evaluation), we see that for the issues COVID-19, defense, and education, there is no difference between politicians with or without a migration background. However, for the issue of migration, we see that if politicians with a migration background compromise -- i.e. voting in favor of the bill proposing to put a stop to migration to Western countries -- they are considered better political leaders compared to politicians without a migration background.
- Regarding how well people feel represented, there is no difference between politicians with and without a migration background.
- Looking at how favorable people feel towards a politcian, we see that people feel more favorable towards politician with a migration background when they compromise on issues related to immigration and education compared to politicians without a migration background, but not for issues related to COVID-19 and defense.
- For career perspectives, people rate the career prospects more negative when a politician with a migration background compromises on the issue migration. For the other issues, there is no difference between politicians with and without a migration background.

\newpage 
\blandscape 
```{r results-main, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results}Results"}
include_graphics(here("report/figures", "save-patchwork-results-1.png"))
```
\elandscape

**_Intersection Hypothesis_** Hypothesis 3a states that women politician with a migration background are evaluated lower when striking a political compromise compared to men politicians without a migration background.
This hypothesis is tested in the lower-left panel of Figure \ref{fig:results} for the separate issues, and in the lower-center-right panel of Figure \ref{fig:results} pooling the issues.

- Trait: no differences for COVID-19 and immigration, insignificant positive effects for defense and education, pooling significant positive effect
- Representation: Positive and signicant for education, nothing for the other issues, pooling significant positive effect
- Favorability:  Positive and signicant for education, nothing for the other issues, pooling significant positive effect
- Career prospects: negative and significant for immigration, nothing for the other issues or in the pooled analysis.

Hypothesis 3b states that women politician with a migration background are evaluated lower when not striking a political compromise compared to men politicians without a migration background.
This hypothesis is tested in the lower-center-left panel of Figure \ref{fig:results} for the separate issues, and in the lower-right panel of Figure \ref{fig:results} pooling the issues.

- Trait: positive and significant for COVID-19, nothing for the other issues or in the pooled analysis.
- Representation: no differences
- Favorability: positive and significant for COVID-19, nothing for the other issues or in the pooled analysis.
- Career prospects:negative and significant for defense,  nothing for the other issues or in the pooled analysis.

Paragraph on what it means for theory, and justification for explorations (below).

## Exploration
To explore hetereogeneous effects, I conduct pooled analyses.

**_Ideology_**

- H1a: People with left-wing ideological position (i.e. `value rile-self $<5$`) have higher evaluations for female politicians who compromise compared to male politicians who compromise. Compared to male politicians, they find women politicians who compromise more favorable, feel better represented and evaluate their leadershipskills higher (yet very small effect). For people with a right-wing ideological position (i.e. `value rile-self $>5$`), there is no statistical significant difference between female and male politicians who compromise.
- H1b: The opposite holds for people with a right-wing ideological position when evaluating women who do not compromise, compared to male politicians: women who stick to their guns are considered more favorable and better leaders than men who stick to their guns. For people with a left-wing ideological position, there is no statistical significant difference between female and male politicians who do not compromise.
- H2: Left-wing more favorable to people politicians with a migration background who compromise than towards politicians without a migration backgroound -- no effect for other three measures. For people with a right-wing ideological position, there is no statistical significant difference between politicians with and without a migration background who compromise. Looking at NOT compromising, we see remarkable differences between people with a left-wing and right-wing ideology. When politicians with a migration background not compromise, they're considered more favorable to left-wingers, but less favorable to right-wingers. Also, not compromising makes them have better leaders according to the left, and worse leaders according to the right.
- H3a: Women with a migration background (compared to women without, and compared to men with and without) who compromise are considered more favorable, better representatives, and to have better career prospects by the left. No statistical difference for the right.
- H3b: No statistical difference for the left of the right

CHECK for DISCUSSION: left-wing over all more agreeable?
```{r "ideology-agree",echo=FALSE}
d %>% 
   select(PreT6, PreT7_2, PreT7_6, PreT7_7) %>%
   mutate(`Ideological Position` = if_else(PreT6>5, 
                                           "Right-Wing", 
                                           "Left-Wing")) %>%
   group_by(`Ideological Position`) %>%
   summarise(c1 = cor(PreT6,PreT7_2),
             c2 = cor(PreT6,PreT7_6),
             c3 = cor(PreT6,PreT7_7)) %>% 
   pivot_longer(cols = c1:c3,
                names_to = "Correlation with",
                values_to = "Pearson's R") %>%
   mutate(`Correlation with` = recode(`Correlation with`,
                                      `c1` = "Against their better judgment, politicians promise more than they can deliver.",
                                      `c2` = "What in politics is called `making compromises' is actually just betraying your principles.",
                                      `c3` = "A strong head of government is good for the Netherlands, even if he stretches the rules a bit to get things done.")) %>%
   kbl(booktabs =T, caption = "\\label{tab:cor_agree}Correlation ideology and agreebleness",
    digits = 2) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
   column_spec(1, width = "3cm") %>%
   column_spec(2, width = "10cm") %>%
   column_spec(3, width = "3cm")
```

\newpage 
\blandscape 
```{r results-ideology, out.width = "100%", fig.align = 'center', fig.cap = "\\label{fig:results-ideology}Results: Interaction with Ideology"}
include_graphics(here("report/figures", "explore-ideology-1.png"))
```
\elandscape

**_Willingness to Accept Compromises_**

- H1a: people with low scores on WTAC, i.e. those who consider political compromise betraying your principles (`scores of 1-3` (inversly coded)), are more likely to consider female politicians who compromise favorable (compared to male politicians who compromise). Both people with high and low scores on WTAC consider female politcians who compromise to be better representatives than male politicians who compromise. People willing to accept compromise consider the career prosects of women who compromise more positive compared to male politicians who compromise.
- H1b: People who are against political compromise rate female politicians who stick to their guns higher in terms of favorability than male politicians who do so. Both people with high and low scores on WTAC consider female politcians who compromise to be better representatives than male politicians who compromise. People willing to accept compromise consider the career prosects of women who compromise more positive compared to male politicians who compromise, also they consider female politicians who do not compromise less favorable, but better leaders.
- H2: People who are against political compromise rate the career prospects of politicians with a migration background who compromise way more negative compared to politicians without a migration background, they are also slightly more negative about the represenation oof politicians with a migration background compared to those without that background. When you are willing to accept political compromises, you are more positive about politicians with a migration background in terms of career prospects, favorability, representatiion and leadership traits. When politicians with a migration background stick to their guns, both people who are and are not willing to accept compromise evaluate them lower compared to politicians without a migration bachground. Moreover, people against political compromises also find politicians with a migration background who stick to their guns less favorable than politicians without a migration background.
- H3a: Female politicians with a migration background are considered more favorable by people who are willing to accept political compromises.
- H3b: Female politicians with a migration background are considered more favorable by people who are not willing to accept political compromises, and less favorable by those who are willing to acceptt compromise. Also, people against political compromisee evaluate the prospects of female politicians with a migration backgound lower than the prospects of the other groups.

CHECK: do people who are willing to accept compromise have higher trust in poltics?

```{r "wtac-trust", echo=FALSE}
d %>% 
   select(PreT7_6, PreT7_1, PreT7_3, PreT7_4, PreT7_5) %>%
   mutate(`Attitude towards Compromise` = if_else(PreT7_6>3, "Positive", 
                         "Negative")) %>%
   group_by(`Attitude towards Compromise`) %>%
   summarise(c1 = cor(PreT7_6,PreT7_1),
             c2 = cor(PreT7_6,PreT7_3),
             c3 = cor(PreT7_6,PreT7_4),
             c4 = cor(PreT7_6,PreT7_5)) %>% 
   pivot_longer(cols = c1:c4,
                names_to = "Correlation with",
                values_to = "Pearson's R") %>%
   mutate(`Correlation with` = recode(`Correlation with`,
                                      `c1` = "The political parties are only interested in my vote and not in my opinion.",
                                      `c2` = "Most politicians are skilled people who know what they are doing.",
                                      `c3` = "Politicians are able to solve the most important problems.",
                                      `c4` = "I feel well represented by Dutch politicians.")) %>%
   kbl(booktabs =T, caption = "\\label{tab:cor_trust}Correlation WtAC and Other measures of Pol. Eff",
    digits = 2) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
   column_spec(1, width = "3cm") %>%
   column_spec(2, width = "10cm") %>%
   column_spec(3, width = "3cm")
```

\newpage 
\blandscape 
```{r results-wtac, out.width = "100%", fig.align = 'center', fig.cap = "\\label{fig:results-wtac}Results: Interaction with Willingness to Accept Compromise"}
include_graphics(here("report/figures", "explore-wtac-1.png"))
```
\elandscape

**_Attitude towards Women in Politics_**

- H1a: Whether or not you're positive towards women in politics, female politicians who compromise are considered more favorable than male politicians -- the same holds for women who do not compromise. For those negative towards women in politics, they evaluate women politicians who compromise as having better career perspectives, being better representatives, yet having lower leadership skills. For those positive about women in politics, female politicians who compromise are seen as worse representatives than males who compromise. The opposte holds for leadership skills. 
- H1b: For those negative towards women in politics, they evaluate women politicians who do not compromise as having better career perspectives, yet being worse representatives. There is no difference in their leadership skills (effect of 0). For those positive about women in politics, female politicians who do not compromise are seen as better representatives than males who do not compromise. The same holds for leadership skills. 
- H3a: People who have a negative attitude towards women in politics evaluate female polticians with a migration background lower. They rate their career perspectives  and leadershipskills more negative. People with positive attitudes towards women in politics rate female politicians with a migration background who compromise -- relative to other groups -- higher in terms of career prospects, favorability and leadership skills.
- H3b: When female politicians with a migration background do not compromise, they are lower evaluated on their career prospects, but are considered to be better representatives by those with a positive attitude towards women in politics. When female politicians with a migration background do not comprimse, their career perspectives are evaluated lower, but they are considered better representatives by people with positive attitudes towards women in politics.

\newpage 
\blandscape 
```{r results-women-politics, out.width = "100%", fig.align = 'center', fig.cap = "\\label{fig:results-women-politics}Results: Interaction with Attitudes towards Women in Politics"}
include_graphics(here("report/figures", "explore-women-politics-1.png"))
```
\elandscape

**_Position on Political Issues_**

- H1a:
   - COVID-19: When opposing COVID-19 policies, people rare female politicians who compromise as more favorable than male politicans who compromise. People who are infacor of the COVID-19 policies do not evaluate male and female politicians who compromise differently
   - Defense: No difference between positive or negative towards defense, people always rate compromising female politicians as more favorable. When you are negative towards defense policies: find women who compromise better representatives, when positive about defense policies, you evaluate their career prosects more positive.
   - Education: When negative position on education, you rate compromising female politicians higher than compromising male politicians in terms of favorability, as representatives, and in terms of their leadership skills. 
   - Immigratin: When estimaing low percentagees of immigration in the Netherlands, you're more positive towards compromising female politicians in terms of career prospects, favorability and representation.

- H1b: 
   - COVID-19: People who are negative towards COVID-19 policies find female politicians who stick to their guns better representatives and evaluate their leadership skills hiigher compared to male politicians who compromise. People who are positive towards COVID-19 politicies find female politicians who stick to their guns more likeable.
   - Defense: People who are negative towards defense policies find female politicians who stick to their guns more favorable, and think they have better career prospects. People who are positive towards defense politicies find female politicians who stick to their guns better representatives.
   - Education: Position on education has no effect for favorability and leadershipskills. For both hold that women sticking to their guns are higher evaluated than male politicians doing the same. If you're negative towards education, you rate the career prospects of women sticking to their guns higher, and when you're positive towards education, you find women sticking to their guns better representatives. 
   - Immigration: Positive on immigration coincides with

- H2: 
   - COVID-19:
   - Defense:
   - Education:
   - Immigration:

- H3a: 
   - COVID-19:
   - Defense:
   - Education:
   - Immigration:

- H3b: 
   - COVID-19:
   - Defense:
   - Education:
   - Immigration:

\newpage 
\blandscape 
```{r results-covid, out.width = "100%", fig.align = 'center', fig.cap = "\\label{fig:results-covid}Results: Interaction with Position on COVID-19"}
include_graphics(here("report/figures", "explore-covid-1.png"))
```
\elandscape

\newpage 
\blandscape 
```{r results-defense, out.width = "100%", fig.align = 'center', fig.cap = "\\label{fig:results-defense}Results: Interaction with Position on Defense"}
include_graphics(here("report/figures", "explore-defense-1.png"))
```
\elandscape

\newpage 
\blandscape 
```{r results-education, out.width = "100%", fig.align = 'center', fig.cap = "\\label{fig:results-education}Results: Interaction with Position on Education"}
include_graphics(here("report/figures", "explore-education-1.png"))
```
\elandscape

\newpage 
\blandscape 
```{r results-immigration, out.width = "100%", fig.align = 'center', fig.cap = "\\label{fig:results-immigration}Results: Interaction with Position on Immigration"}
include_graphics(here("report/figures", "explore-immigration-1.png"))
```
\elandscape

# Discussion

\newpage
# References