### 2021-05-27 report/figures/table_descriptives.tex
rerun analyses

# Table: Descriptive Information of Variables under Study

<table>
 <thead>
  <tr>
   <th style="text-align:left;"> Variables </th>
   <th style="text-align:right;"> Mean Value </th>
   <th style="text-align:right;"> St. Dev </th>
   <th style="text-align:right;"> Min. Value </th>
   <th style="text-align:right;"> Max. Value </th>
  </tr>
 </thead>
<tbody>
  <tr>
   <td style="text-align:left;"> Treatment: Striking Compromise </td>
   <td style="text-align:right;"> 0.50 </td>
   <td style="text-align:right;"> 0.50 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 1 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Gender </td>
   <td style="text-align:right;"> 0.38 </td>
   <td style="text-align:right;"> 0.49 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 1 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Age </td>
   <td style="text-align:right;"> 33.64 </td>
   <td style="text-align:right;"> 16.37 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 72 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Region </td>
   <td style="text-align:right;"> 2.93 </td>
   <td style="text-align:right;"> 1.39 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Urbanness </td>
   <td style="text-align:right;"> 2.28 </td>
   <td style="text-align:right;"> 1.45 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Migration Background </td>
   <td style="text-align:right;"> 1.30 </td>
   <td style="text-align:right;"> 0.89 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Employment </td>
   <td style="text-align:right;"> 3.36 </td>
   <td style="text-align:right;"> 2.67 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 8 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Income </td>
   <td style="text-align:right;"> 9.75 </td>
   <td style="text-align:right;"> 2.74 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 14 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Education </td>
   <td style="text-align:right;"> 1.27 </td>
   <td style="text-align:right;"> 0.53 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 3 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Treatment: Gender Politician </td>
   <td style="text-align:right;"> 0.50 </td>
   <td style="text-align:right;"> 0.50 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 1 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Treatment: Immigration Background Politician </td>
   <td style="text-align:right;"> 0.50 </td>
   <td style="text-align:right;"> 0.50 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 1 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Treatment: Women Politician with Immigration Background </td>
   <td style="text-align:right;"> 0.25 </td>
   <td style="text-align:right;"> 0.43 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 1 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Immigration (%) </td>
   <td style="text-align:right;"> 22.76 </td>
   <td style="text-align:right;"> 12.78 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 100 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Position: Immigration </td>
   <td style="text-align:right;"> 5.88 </td>
   <td style="text-align:right;"> 2.04 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 10 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Position: Defense </td>
   <td style="text-align:right;"> 2.50 </td>
   <td style="text-align:right;"> 1.05 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 4 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Position: Education </td>
   <td style="text-align:right;"> 3.12 </td>
   <td style="text-align:right;"> 0.82 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 4 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Attitude: Women in Politics </td>
   <td style="text-align:right;"> 3.41 </td>
   <td style="text-align:right;"> 0.92 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Ideology </td>
   <td style="text-align:right;"> 4.46 </td>
   <td style="text-align:right;"> 2.20 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 10 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Political Efficacy (1) </td>
   <td style="text-align:right;"> 3.12 </td>
   <td style="text-align:right;"> 1.03 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Political Efficacy (2) </td>
   <td style="text-align:right;"> 3.94 </td>
   <td style="text-align:right;"> 0.85 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Political Efficacy (3) </td>
   <td style="text-align:right;"> 3.16 </td>
   <td style="text-align:right;"> 0.95 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Political Efficacy (4) </td>
   <td style="text-align:right;"> 2.87 </td>
   <td style="text-align:right;"> 0.97 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Political Efficacy (5) </td>
   <td style="text-align:right;"> 2.83 </td>
   <td style="text-align:right;"> 1.00 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Political Efficacy (6) </td>
   <td style="text-align:right;"> 2.39 </td>
   <td style="text-align:right;"> 1.04 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Political Efficacy (7) </td>
   <td style="text-align:right;"> 2.74 </td>
   <td style="text-align:right;"> 1.19 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Position: COVID-19 </td>
   <td style="text-align:right;"> 2.86 </td>
   <td style="text-align:right;"> 1.18 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> DV: Trait Evaluation </td>
   <td style="text-align:right;"> 3.21 </td>
   <td style="text-align:right;"> 0.64 </td>
   <td style="text-align:right;"> 1 </td>
   <td style="text-align:right;"> 5 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> DV: Favorability </td>
   <td style="text-align:right;"> 5.41 </td>
   <td style="text-align:right;"> 2.01 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 10 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> DV: Representation </td>
   <td style="text-align:right;"> 5.86 </td>
   <td style="text-align:right;"> 1.77 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 10 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> DV: Career Prospects </td>
   <td style="text-align:right;"> 4.64 </td>
   <td style="text-align:right;"> 1.97 </td>
   <td style="text-align:right;"> 0 </td>
   <td style="text-align:right;"> 10 </td>
  </tr>
</tbody>
</table>